---
title: "BDCD Business Analytics Lab Exercise"
author: "Prof Di Cook, Steph Kobakian, Stuart Lee, Nick Spyrison"
date: "Econometrics & Bus Stat, Monash, Clayton campus, 22/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  error = FALSE)
```

## Links to reading

https://compare.energy.vic.gov.au

http://www.smartmeters.vic.gov.au/#

[https://www.computerworld.com](https://www.computerworld.com/article/2497143/business-intelligence/business-intelligence-beginner-s-guide-to-r-introduction.html)

[https://github.com/Monash-BDCD/energy](https://github.com/Monash-BDCD/energy)

## Exercise 1: Background work

Read in Di's energy data. Look at the format of the data, and then rearrange it to a tidier format.

```{r readdata, fig.width=3, fig.height=7, fig.fullwidth=TRUE}
library(tidyverse)
library(lubridate)
library(tsibble)
library(sugrrants)
library(glue)
library(emo)
elec <- read_csv("data/di.csv", skip=1,
                 col_names = c("id", "date", paste0("d", 1:48), paste0("stuff", 1:5)),
                 col_types = "ccddddddddddddddddddddddddddddddddddddddddddddddddccccc")
elec %>% select(id:d5) %>% head(3)
```

Here's the wrangling, `r set.seed(2016); emo::ji("workout")`, and new format:

```{r wrangle}
vic_holidays <- holiday_aus(2017:2019, state = "VIC")
elec <- elec %>% filter(id == 300)
elec <- elec %>%
  mutate(date = ymd(date)) %>%
  select(id:d48) %>%
  gather(halfhour, kwh, d1:d48) %>%
  mutate(halfhour = as.numeric(sub("d", "", halfhour))/2) %>%
  arrange(date, halfhour) %>%
  mutate(wday = wday(date, label = TRUE, abbr = TRUE,
                     week_start = 1),
         month = month(date, label = TRUE, abbr = TRUE),
         year = year(date)) %>%
  mutate(dt = ymd_hm(glue("{date} 12:00"),
                     tz = "Australia/Melbourne") +
           minutes(60*halfhour)) %>% 
  mutate(work = ifelse(wday %in% c("Mon", "Tue", "Wed", "Thu", "Fri"), "workday", "holiday")) %>%
  mutate(work = ifelse(date %in% vic_holidays$date, "holiday", work))
elec %>% head(3) 
```

## Exercise 2: Plot (some of) the data in a calendar layout


```{r fig.width=3.5, fig.height=3.5, echo=TRUE, tidy=TRUE}
p1 <- elec %>%
  filter(date < dmy("01022019"), date >= dmy("01122018")) %>%
  frame_calendar(x = halfhour, y = kwh, date = date, ncol = 4) %>%
  ggplot(aes(x = .halfhour, y = .kwh, group = date, colour=factor(work))) +
  geom_line() +
  scale_colour_brewer("work", palette = "Dark2") +
  theme(legend.position="none")
prettify(p1)
```

## Exercise 3: Combine with weather data

```{r getweather}
library(bomrang)
library(viridis)
stations <- sweep_for_stations(latlon = c(-37.8136, 144.9631)) #Melbourne lat/long is 
maxtemp <- get_historical(stationid = "086282", type = "max") 
maxtemp <- maxtemp %>%  
  mutate(date = paste(maxtemp$Year, maxtemp$Month, maxtemp$Day, sep="-")) %>%
  mutate(date = ymd(date))
maxtemp_DF <- maxtemp %>% filter(Year > 2017, Month %in% c(12, 1)) 
elec <- elec %>%   
  left_join(maxtemp_DF) 
```

```{r fig.width=3.5, fig.height=4, echo=TRUE, tidy=TRUE}
p1 <- elec %>%
  filter(date < dmy("01022019"), date >= dmy("01122018")) %>%
  frame_calendar(x = halfhour, y = kwh, date = date, ncol = 4) %>%
  ggplot(aes(x = .halfhour, y = .kwh, group = date, colour=Max_temperature)) +
  geom_line() +
  scale_colour_viridis_c("temperature", option="inferno", direction=-1) +
  theme(legend.position = "bottom")
prettify(p1)
```

## Exercise 4: Make a simple plot, and add some interactivity

```{r fig.width=3, fig.height=3, echo=TRUE}
library(plotly)
p <- elec %>%
  filter(year(date) == 2018) %>%
  group_by(date) %>%
  summarise(kwh = sum(kwh)) %>%
  left_join(maxtemp_DF) %>%
  ggplot(aes(x=Max_temperature, y=kwh, text=date)) + geom_point()
ggplotly(p)
```

## Exercise 6: Compare day of the week

```{r fig.width=4, fig.height=3, echo=TRUE}
p <- elec %>%
  filter(year(date) == 2018) %>%
  group_by(date) %>%
  summarise(kwh = sum(kwh)) %>%
  mutate(wday = wday(date, label = TRUE, week_start = 1)) %>%
  ggplot(aes(x=wday, y=kwh)) + geom_boxplot()
ggplotly(p)
```

## Your turn

1. Easy tasks: 
    a. Create a test Rmarkdown document
    b. Compile it into an html document
2. Medium task: Create a plot to answer these questions: 
    a. What day of the week does the household use more energy? What day of the week is there more variation in energy use?
    b. What time of day does the household wake up? And go to sleep? Is it different on a holiday vs a working day?
    c. Can you guess when Di's son was home from Uni? Was there a time when the family was away on holidays?
3. Difficult: Make a new app to study energy usage. The steps to do this are:
    a. Run the sample app, and play with changing dates
    b. Add a numerical summary to the output  
    c. Make a new tab with a new display, and numerical summary, something that you think is useful to track in energy usage.

## How to win the grand prize

There is a prize for the team that works together the best on the lab activities. 

1. Get the best score in the interactive online quiz (after lunch)
2. Produce the most creative web app, with interesting data plots, that someone can use to explore their energy use



