---
title: "Check your learning on the BDCD 2019 Business Analytics Challenge"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Quiz questions


```{r quiz}
quiz(#1
  question("When were smart meters rolled out in Victorian households?",
    answer("2011 to 2015"),
    answer("2012 to 2016"),
    answer("2013 to 2016", correct = TRUE),
    answer("2014 to 2018")
  ),#2
  question("How many smart meters have been installed?",
    answer("2.8 million", correct = TRUE),
    answer("3.2 million"),
    answer("3.5 million"),
    answer("20 million")
  ),#3
  question("How much money does the Victorian Government give you to compare your energy costs with different providers?",
    answer("None"),
    answer("$15"),
    answer("$50", correct = TRUE),
    answer("$500")
  ),#4
  question("Why would you use the R language for data analysis?",
    answer("more than 10,000 add-on package", correct = TRUE),
    answer("95,000+ members of LinkedIn's R group", correct = TRUE),
    answer("more than 400 R Meetup groups", correct = TRUE),
    answer("you're standing on the shoulders of giants", correct = TRUE)
  ),#5
  question("Rmarkdown is?",
    answer("a file format for making dynamic documents with R", correct = TRUE),
    answer("an easy-to-write plain text format", correct = TRUE),
    answer("a document with chunks of embedded R code", correct = TRUE),
    answer("an advanced Excel sheet")
  ),#6
  question("The three back-ticks ``` indicate",
    answer("the start of a quote"),
    answer("the equivalent of an exclamation mark"),
    answer("the start of a block of code in an Rmarkdown document", correct = TRUE),
    answer("the start of instructions to make a data plot")
  ),#7
  question("To make a plot of data using a grammar of graphics in R, which **package** should you use?",
    answer("plot"),
    answer("plotly"),
    answer("ggplot", correct = TRUE),
    answer("sugrrants")
  ),#8
  question("To make a calendar plot of data in R, which **package** should you use?",
    answer("plot"),
    answer("plotly"),
    answer("ggplot"),
    answer("sugrrants", correct = TRUE)
  ),#9
  question("To make an interactive plot of data in R, which **package** should you use?",
    answer("plot"),
    answer("plotly", correct = TRUE),
    answer("ggplot"),
    answer("sugrrants")
  ),#10
  question("What type of plot does this code make? ggplot(aes(x = Max_temperature, y = kwh, text = date)) + geom_point()",
    answer("scatterplot", correct = TRUE),
    answer("boxplot"),
    answer("calendar plot"),
    answer("point plot")
  ),#11
  question("What does the function *filter* do",
    answer("selects columns of data"),
    answer("selects rows of data", correct = TRUE),
    answer("transforms a column in the data"),
    answer("cleans the data")
  ),#12
  question("What does the function *mutate* do",
    answer("selects columns of data"),
    answer("selects rows of data"),
    answer("transforms a column in the data", correct = TRUE),
    answer("cleans the data")
  ),#13
  question("Professor Cook's energy use in January 2019 is _______ January 2018?",
    answer("much lower than", correct = TRUE),
    answer("much higher than"),
    answer("about equal to"),
    answer("I have no idea")
  ),#14
  question("Which day of the week has the highest median usage?",
    answer("Monday"),
    answer("Tuesday"),
    answer("Wednesday"),
    answer("Thursday"),
    answer("Friday"),
    answer("Saturday"),
    answer("Sunday", correct = TRUE)
  ),#15
  question("What is Professor Cook's vision for Melbourne 2050? A city where ",
    answer("people use energy efficiently, and live in a green, clean air environment"),
    answer("everyone knows your name, and cares about who you are, and that you stay healthy and happy"),
    answer("the information about your energy usage is clearly shown to you by your energy provider"),
    answer("people have control of data about them. They can download this data, and make plots to learn about their life, and what they might do to improve it.", correct = TRUE)
  )
)
```

